---
title: "Hospital Exploratory Analysis"
output: html_notebook
author: Nicole Wright
---

OAITI has access to all USA Hospital data. The goal of this exploratory analysis is to find intersting questions to ask about the data.  


```{r Load Packages}
library(ggplot2)
library(plyr)
library(tidyverse)
library(readr)
library(tidycensus)
```

```{r Read-in Data}
hospital <- read.csv("C:/Users/Nicole/Desktop/OAITI/Hospital Data/Hospital.csv")
```

```{r First look}
View(hospital)
head(hospital)
glimpse(hospital)
```

```{r hospital data clean up}
#examine footnotes
hospital %>%
  select(ï..Provider.ID, contains("footnote")) 
#Remove footnotes, rename columns

H <- hospital %>%
  select(-contains("footnote"), -Phone.Number, -Efficient.use.of.medical.imaging.national.comparison) %>%
  filter(State != "PR") %>%
  filter(State != "AS") %>%
  filter(State != "MP") %>%
  filter(State != "GU") %>%
  filter(State != "VI") 

  names(H) <- c("ID", "Name", "Address", "City", "State", "Zip_Code", "County", "Type", 
             "Ownership", "Emergency_Services", "Meaningful_EHRs", "Overall_Rating", "Mortality", "Safety", "Readmission", "Patient_experience", "Effectiveness", "Timeliness") 
 
glimpse(H)
View(H)
```


2010 US Census data, using the tidycensus package ( 1/1/2019).
Key: 6bfbcee4b4c37a9683ac9c0e1d174d5f0ad67674
pdf: https://cran.r-project.org/web/packages/tidycensus/tidycensus.pdf
```{r 2010 Census set up}
census_api_key("6bfbcee4b4c37a9683ac9c0e1d174d5f0ad67674", install = TRUE, overwrite = TRUE)

# First time, reload your environment so you can use the key without restarting R.
readRenviron("~/.Renviron")
```


Variables of Interest from Hospital data: Provider.ID, Hospital.Name, City, State, County.Name, Hospital.Type, Hospital.Ownership, Hospital Overall.rating, all comparisons (?)

variables of Interest from census data: by county - population, Income, race, age (?)

Questions:
How is the frequency of hospitals related to county population, Income, or race?
Is there any correlation between hospital rating and population, Income, or race?

```{r checking out variables}
#2010 Census, population and race
LV1 <- load_variables(2010, "sf1", cache = TRUE) 
View(LV1)
VarsC <- c("White" = "P003002", "Black" = "P003003", "Native_American" = "P003004", "Asian" = "P003005", "Pacific_Islander" = "P003006", "Other" = "P003007", "Two_or_more" = "P003008", "Hispanic/Latino" = "P004003", "Median Age" = "P013001")

#2010 ACS, income data
LV2<- load_variables(2010, "acs5", cache = TRUE)
View(LV2)
VarsA <- c("Median Family Income" = "B19113_001", "Income per Capita" = "B19301_001")

```

Upon examining Median Family Income versus Income per Captia I've noticed that Median income is much higher than per capita income. Is this because per capita income is counting all the non-working people in the average? Moving forward I will use Median Family Income

```{r get the data}
a1 <- get_decennial(geography = "county", variables = VarsC, summary_var = c(Total_Population = "P001001", Median_Age = "P013001"), year = 2010, geometry = TRUE, cache_table = TRUE, shift_geo = TRUE)
View(a1)

a2 <- get_acs(geography = "county", variables = c("Median Family Income" = "B19113_001"), year = 2010, geometry = TRUE, cache_table = TRUE)
View(a2)

```
```{r Editing a1 and a2 data sets}
a1 <- a1 %>%
  rename(c("NAME" ="County", "variable" ="Race", "value" = "Population")) %>%
  mutate(per_Pop = Population/Total_Population * 100)

State <- unlist(strsplit(a1$County, split = ",")) %>%
  matrix(ncol = 2, byrow = TRUE) %>% 
  as.tibble() %>%
  rename(c("V1" = "County", "V2" = "State"))

a1$State <- State$State


a1_final <- a1 %>%
   select(GEOID, County, State, Median_Age, Race, per_Pop, Total_Population, geometry_census = geometry)
View(a1_final)

a2_final <- a2 %>%
  as.tibble() %>%
  rename(c("NAME" ="County", "estimate" = "Median Family Income")) %>%
  select(-variable) 
View(a2_final)
```


```{r Joining a1 and a2 into census data set}
census <- left_join(a1_final, a2_final, by = "GEOID")

census <- census %>%
  select(-County.y) 
census$County <- census$County.x 
```
 
##Exploratory Analysis

```{r Explore H data}
H %>% 
  group_by(State) %>%
  tally() %>%
  plot()

tally <- H %>% 
  group_by(State) %>%
  tally()

tally %>%
  arrange(n) 
count(tally$State)
```
*Turns out there are 5 US territories included here: PR, AS, MP, GU, VI
* Plot of state hospitals tallied show a pretty even distribution among states with two very high outliers,o TX(405), and CA (339). 
