---
title: "Hospital Exploratory Analysis"
output: html_notebook
author: Nicole Wright
---

OAITI has access to all USA Hospital data. The goal of this exploratory analysis is to find intersting questions to ask about the data.  


```{r Load Packages}
library(ggplot2)
library(dplyr)
library(tidyverse)
library(readr)
library(tidycensus)
library(tmap)
library(tmaptools)
library(sf)
```

```{r Read-in Data}
hospital <- read.csv("C:/Users/Nicole/Desktop/OAITI/Hospital Data/Hospital.csv")
```

```{r First look}
View(hospital)
head(hospital)
glimpse(hospital)
```

```{r hospital data clean up}
#examine footnotes
hospital %>%
  select(ï..Provider.ID, contains("footnote")) 
#Remove footnotes, rename columns

H <- hospital %>%
  select(-contains("footnote"), -Phone.Number, -Efficient.use.of.medical.imaging.national.comparison) %>%
  filter(State != "PR") %>%
  filter(State != "AS") %>%
  filter(State != "MP") %>%
  filter(State != "GU") %>%
  filter(State != "VI") 

  names(H) <- c("ID", "Name", "Address", "City", "State", "Zip_Code", "County", "Type", 
             "Ownership", "Emergency_Services", "Meaningful_EHRs", "Overall_Rating", "Mortality", "Safety", "Readmission", "Patient_experience", "Effectiveness", "Timeliness") 
 
glimpse(H)
View(H)
```


2010 US Census data, using the tidycensus package ( 1/1/2019).
Key: 6bfbcee4b4c37a9683ac9c0e1d174d5f0ad67674
pdf: https://cran.r-project.org/web/packages/tidycensus/tidycensus.pdf
```{r 2010 Census set up}
census_api_key("6bfbcee4b4c37a9683ac9c0e1d174d5f0ad67674", install = TRUE, overwrite = TRUE)

# First time, reload your environment so you can use the key without restarting R.
readRenviron("~/.Renviron")
```


Variables of Interest from Hospital data: Provider.ID, Hospital.Name, City, State, County.Name, Hospital.Type, Hospital.Ownership, Hospital Overall.rating, all comparisons (?)

variables of Interest from census data: by county - population, Income, race, age (?)

Questions:
How is the frequency of hospitals related to county population, Income, or race?
Is there any correlation between hospital rating and population, Income, or race?

```{r checking out variables}
#2010 Census, population and race
LV1 <- load_variables(2010, "sf1", cache = TRUE) 
View(LV1)
VarsC <- c("White" = "P003002", "Black" = "P003003", "Native_American" = "P003004", "Asian" = "P003005", "Pacific_Islander" = "P003006", "Other" = "P003007", "Two_or_more" = "P003008", "Hispanic/Latino" = "P004003", "Median Age" = "P013001")

#2010 ACS, income data
LV2<- load_variables(2010, "acs5", cache = TRUE)
View(LV2)
VarsA <- c("Median Family Income" = "B19113_001", "Income per Capita" = "B19301_001")

```

Upon examining Median Family Income versus Income per Captia I've noticed that Median income is much higher than per capita income. Is this because per capita income is counting all the non-working people in the average? Moving forward I will use Median Family Income

```{r get the data}
a1 <- get_decennial(geography = "county", variables = VarsC, summary_var = c(Total_Population = "P001001", Median_Age = "P013001"), year = 2010, geometry = TRUE, cache_table = TRUE, shift_geo = TRUE)

a2 <- get_acs(geography = "county", variables = c("Median Family Income" = "B19113_001"), year = 2010, geometry = TRUE, cache_table = TRUE)
```

```{r Editing a1 and a2 data sets}
VarsC <- c("White" = "P003002", "Black" = "P003003", "Native_American" = "P003004", "Asian" = "P003005", "Pacific_Islander" = "P003006", "Other" = "P003007", "Two_or_more" = "P003008", "Hispanic/Latino" = "P004003", "Median Age" = "P013001")

a1 <- get_decennial(geography = "county", variables = VarsC, summary_var = c(Total_Population = "P001001", Median_Age = "P013001"), year = 2010, geometry = TRUE, cache_table = TRUE, shift_geo = TRUE, key = "ab33f4101bd279734ec43e7d7b9a12ace24d8eeb")

a1_edit <- a1 %>%
    rename(County = NAME, Race = variable, Population = value) %>%
    mutate(per_Pop = Population/Total_Population * 100)

State <- unlist(strsplit(a1_edit$County, split = ", ")) %>%
    matrix(ncol = 2, byrow = TRUE) %>% 
    as_tibble() %>%
    rename(County = V1, State = V2)

a1_edit$State <- State$State

a1_final <- a1_edit %>%
    as_tibble() %>%
    select(GEOID, County, State, Median_Age, Race, per_Pop, Total_Population, geometry_a1 = geometry)

a2_final <- a2 %>%
  as.tibble() %>%
  rename(Median_Family_Income = estimate, geometry_a2 = geometry) %>%
  select(-NAME, -variable) 
```


```{r Joining a1 and a2 into census data set}
census <- left_join(a1_final, a2_final, by = "GEOID")
  

```
 
##Exploratory Analysis

```{r Explore H data}
tally <- H %>% 
  group_by(State) %>%
  tally()

plot(tally)
summary(tally)

tally %>%
  arrange(n) 

sub1 <- census %>%
  group_by(State) %>%
  select(Total_Population, Median Family Income, Median_Age)
```
*Turns out there are 5 US territories included here: PR, AS, MP, GU, VI in the Hospital data
* Plot of state hospitals tallied show a pretty even distribution among states with two very high outliers,o TX(405), and CA (339). The lowest number of hospital per state makes sense for DE, DC, and RI, these areas have the smallest area.
* http://beef2live.com/story-ranking-states-area-89-118259, get land mass area (square miles)

```{r US Map plotting}
#attempt 1
census_sf <-st_sf(census)
class(census_sf)

US_map2 <- tm_shape(census_sf) +  tm_polygons("Median_Age", palette = "-Blues", contrast = .7, id = "County", title = "Median Age by County") 

US_map2
#produces an ok looking map, takes FOREVER to run
```

*First Focus*

First I will analyze the census data with a focus on Iowa census data. I may add additions variables as analysis continues. I will be using a combination of tmap and ggplot until I decided which package is the best for my use.


```{r Iowa subset - tmap}
#subset by Iowa, determine if load time decreases - use tmap
census %>%
  filter(State == "Iowa") %>%
  st_sf() %>%
  tm_shape() + 
  tm_polygons("Median_Age", palette = "-Blues", contrast = .7, 
              id = "County", title = "Median Age by County")

census_IA <- census %>% 
  filter(State == "Iowa") %>%
  rename(geometry = geometry_a1) %>%
  select(-geometry_a2)

census_IA %>%
  arrange(desc(Median_Age)) %>%
  head()
#Highest median age is 48.1
```

```{r Iowa subset - ggplot}

ggplot(data = census_IA) +
  geom_sf(aes(fill = Median_Age))
```


```{r Iowa Median Family Income analysis}
census_IA$IncomeR <- round(census_IA$Median_Family_Income, -4)

ggplot(data = census_IA) +
  geom_sf(aes(fill = factor(IncomeR))) + 
  scale_fill_brewer(palette = "Set2") +
  labs(title = "Iowan Median Family Income") + 
  coord_sf(datum = NA) + 
  theme(plot.background = element_rect(color = "white"), legend.title = element_blank()) 

#theme_minimal and theme_void do not get rid of lines... no does specifying 
#Add dollar signs to legend. Do we need legend title?
#adjust tilt
census_IA %>% 
  filter(Median_Family_Income == max(Median_Family_Income) | Median_Family_Income == min(Median_Family_Income))

census_IA %>%
  mutate(ave = mean(Median_Family_Income), med = median(Median_Family_Income))

```

Examing the Iowa median family income by county we can see the majority of counties make either 50,000 or 60,000 dollars. Appanoose County has the lowest median family income at  $41,250 dollars and Dallas county has the largest median family income at 84,018. The average median family income is 58,233.48 and a median income amoung counties of 57,208. The mean and median of median family incomes amount counties in Iowa are close together, which suggests that there is not much skew in the data. 

```{r Discretize numerical}
#Using cut_interval

census_IA$Age <- cut_interval(census_IA$Median_Age, n = 4, length = NULL, labels = c("25-30", "30-35", "35-40", ">40"))
#works like a charm!

```


```{r plot discretize Median Age}
ggplot(data = census_IA) +
  geom_sf(aes(fill = Age)) + 
  scale_fill_brewer(palette = "Set2") +
  labs(title = "Iowan Median Age") + 
  coord_sf(datum = NA) + 
  theme(plot.background = element_rect(color = "white"), legend.title = element_blank())


```

```{r plot total population}
# soltuionsImperfect
census_IA$Tot_Pop <- cut_interval(census_IA$Total_Population, n = 20, length = NULL
census_IA$Tot_Pop <- round(census_IA$Total_Population, -4)
census_IA$Tot_Pop <- log10(census_IA$Total_Population)


ggplot(data = census_IA) +
  geom_sf(aes(fill = factor(Tot_Pop))) + 
  labs(title = "Iowan population per county") + 
  coord_sf(datum = NA) + 
  theme(plot.background = element_rect(color = "white"), legend.title = element_blank())

```

