---
title: "Hospital Exploratory Analysis"
output: html_notebook
author: Nicole Wright
---

OAITI has access to all USA Hospital data. The goal of this exploratory analysis is to find intersting questions to ask about the data.  


```{r Load Packages}
library(ggplot2)
library(tidyverse)
library(readr)
library(tidycensus)
```

```{r Read-in Data}
hospital <- read.csv("C:/Users/Nicole/Desktop/OAITI/Hospital Data/Hospital.csv")
```

```{r First look}
View(hospital)
head(hospital)
glimpse(hospital)
```

2010 US Census data, using the tidycensus package ( 1/1/2019).
Key: 6bfbcee4b4c37a9683ac9c0e1d174d5f0ad67674
pdf: https://cran.r-project.org/web/packages/tidycensus/tidycensus.pdf
```{r 2010 Census set up}
census_api_key("6bfbcee4b4c37a9683ac9c0e1d174d5f0ad67674", install = TRUE, overwrite = TRUE)

# First time, reload your environment so you can use the key without restarting R.
readRenviron("~/.Renviron")
```


Variables of Interest from Hospital data: Provider.ID, Hospital.Name, City, State, County.Name, Hospital.Type, Hospital.Ownership, Hospital Overall.rating, all comparisons (?)

variables of Interest from census data: by county - population, Income, race, age (?)

Questions:
How is the frequency of hospitals related to county population, Income, or race?
Is there any correlation between hospital rating and population, Income, or race?

```{r checking out variables}
LV1 <- load_variables(2010, "sf1", cache = TRUE) 
View(LV1)
VarsC <- c("White" = "P003002", "Black" = "P003003", "Native_American" = "P003004", "Asian" = "P003005", "Pacific_Islander" = "P003006", "Other" = "P003007", "Two_or_more" = "P003008", "Hispanic/Latino" = "P004003")


LV2<- load_variables(2010, "acs5", cache = TRUE)
View(LV2)

```


```{r attempt1}
a1 <- get_decennial(geography = "county", variables = VarsC, summary_var = c(Total_Population = "P001001", Median_Age = "P013001"), year = 2010, geometry = TRUE, cache_table = TRUE, shift_geo = TRUE)
View(a1)

```

